---
title: "Lab FAQ"
---

```{r, warning=FALSE, include=FALSE, message=FALSE}
library(tidyverse)
library(kableExtra)
```

<br><br>

# What do I submit?

You will be creating two files; an Rmd file containing your code and a .html file for viewing your finished document.

You need to submit both of these files on the relevant Canvas assignment page.

Remember you can also add comments to your submission as/if needed on the canvas page, or message Dr G.

<br><br>

# How am I graded?

There are 100 points available for each of the 5 labs. Remember you can drop your lowest lab or exam (e.g. you can skip one)

You can see the SPECIFIC RUBRIC we use for this lab on the canvas page. It's literally how we will award the marks.

We will also provide as much feedback as possible, so please remember to click the rubric button to take a look.

<br><br>

# What your grade means

Overall, here is what your lab should correspond to:

```{r, echo=FALSE}
rubric <- readxl::read_excel("in_G364_Lab0Rubric.xlsx")
knitr::kable(rubric) %>%   
  kable_classic_2() %>%
  kable_styling(bootstrap_options = c("striped", "hover", "responsive"))


```

<br><br>

# What's the late policy?

If you have technical problems & have sent me canvas message with our code and a screenshot of the issue BEFORE THE DEADLINE, all late penalties go away until we have fixed it.

**Submitting late is a risk.** The lab might be graded at any time past the deadline with no notice given. If worked answers are released then that is too late.

But.. I would prefer you submit late than not at all. If you're going to be late for any reason (you don't need to explain why), send Dr G a canvas message.

**I will only start penalizing if you're consistently late and I have sent a written warning.** See the course syllabus for more info. <https://psu.instructure.com/courses/2243429/assignments/syllabus>

<br><br>

# Can I work with friends/my lab partner? What counts as cheating?

You can talk with each other to help during these labs e.g. you can see each others screens and work out answers together. I will try as much as possible to provide alternate datasets.

**BUT NEVER SHARE CODE/SEND SCRIPTS. ALL TEXT SHOULD BE IN YOUR OWN WORDS.**

For example:

-   Working out together the right way to apply the `seq()` function, then each independently adding it to your own scripts is OK.

-   **Copy/pasting your friend's code/text is not.**.

-   So DISCUSS with friends, but then turn away to your own screen and write your own code and text.

I will be randomly running lab reports through plagiarism software.
