---
title: "Tutorial 7: Summarising Data"
output:
  html_document:
    toc: true
    toc_float: yes
    toc_depth: 2
    df_print: paged
    theme: flatly
    highlight: tango
---

```{r setup, include=FALSE}
# OPTIONS -----------------------------------------------
knitr::opts_chunk$set(echo = TRUE, 
                      warning=FALSE, 
                      message = FALSE)
```

```{r, include=FALSE}
# PACKAGES-----------------------------------------------
# Tutorial packages
library(vembedr)
library(skimr)
library(yarrr)
library(RColorBrewer)
library(GGally) 
library(tidyverse)
library(plotly)
library(readxl)
library(rvest)
library(biscale)
library(tidycensus)
library(cowplot)
library(units)
```

<br><br>

This tutorial covers loads of summary statistics that might be useful.

<br>

# 1. General pointers

Here's the structure of EVERY command in R.

```{r, eval=FALSE}

output_variable  <- COMMAND( input_variable , other_options)

```

-   **`COMMAND`** is the function used to read the data
    -   e.g.,**`read.csv`** for CSV files <br><br>
-   **`input_variable"`** should be replaced with the case sensitive R variable you want to apply it to. Look at your environment tab to see what you have saved. <br><br>
-   **`other_options`** includes (optional) additional options for your
    command
    -   e.g. nrow specifies the number of rows within the read.csv()
        command <br><br>
-   **`"output_variable"`** is the name of the variable you
    choose to assign your data to.
    -   e.g `x <- 5` assigns the number 5 to a variable called x <br>
        Then x <- 5+x, will first calculate the answer (5+5), then 
        OVERWRITE the x with the answer e.g x = 10
         



<br><br>

# 7. Summarising data

Here I will show a few examples for the a non spatial dataset on houseprices.

First I load the data here

```{r,warning=FALSE,message=FALSE}
data("HousesNY", package = "Stat2Data")
```

<br><br>

## 7.1. Looking at the data itself 

To have a look at the data there are many options. You can:

-   click on its name in the environment tab
-   Type its name into the console or into a code chunk (e.g. for our table, type `piratedataset` into the console or a code chunk)
-   Run the command `View(variable_name)` (View is a command from the tidyverse package).<br> This will open the data in a new tab. DON'T PUT THIS IN A CODE CHUNK.
-   Run the command `head(variable_name)` to see the first 6 lines or so (good for quick checks)
-   Run the command `glimpse(variable_name)` to get a nice summary.
-   Run the command `names(variable_name)` to get the column names.

DO NOT PUT View(dataname) into a code chunk (or remove it before you knit). It breaks R-studio sometimes

<br>

For example

```{r, eval=FALSE}
# Note, there are sometimes more columns to the right, use the arrow to see
head(HousesNY)
```

To see what the column names are, you can use the `names(dataset)` command. 
I use this in the console A LOT for copy/pasting names into my code/report.

```{r}
names(HousesNY)
```

Or the glimpse command:

```{r, eval=FALSE}
glimpse(HousesNY)
```

<br> <br>

## 7.2. Number of rows and columns

To find the number of rows and columns, these are useful. Or look at the environment tab, or some summaries include it

```{r, eval=FALSE}
nrow(HousesNY)
ncol(HousesNY)
```


```{r, eval=FALSE}
#or both dimensions
dim(HousesNY)
```

## 7.3. Finding the type of data

To see what type of data R *thinks* you have, try the class command

```{r, eval=FALSE}
class(HousesNY)
```

or for a column

```{r, eval=FALSE}
class(HousesNY$Price)
```


<br><br>

## 7.4.  Single column statistics

Or you can do things manually, using the \$ symbol to choose a column. All of this is for the price column

```{r, eval=FALSE}
mean(HousesNY$Price)
median(HousesNY$Price)
mode(HousesNY$Price)
sd(HousesNY$Price)
var(HousesNY$Price)
IQR(HousesNY$Price)
range(HousesNY$Price)
```


<br><br>

## 7.5.  Useful summaries

To look at the summaries there are a load of options. Choose your favourites:

-   `summary(dataset)`
-   `skim(dataset)` in the skimr package
-   `summarize(dataset)` in the papeR package. This looks pretty powerful, I'm just learning it

None are better or worse than others - simply choose what works for you in the moment.

```{r, eval=FALSE}
summary(HousesNY)
```

```{r, eval=FALSE}
library(skimr) # you would need to install this
skim(HousesNY)
```

```{r, eval=FALSE}
library(pillar) # you would need to install this
glimpse(HousesNY)
```

or

```{r, eval=FALSE}
str(HousesNY)
```




