---
title: "Tutorial 5: Reading In Data"
output:
  html_document:
    toc: true
    toc_float: yes
    toc_depth: 2
    df_print: paged
    theme: flatly
    highlight: tango
---

```{r setup, include=FALSE}
# OPTIONS -----------------------------------------------
knitr::opts_chunk$set(echo = TRUE, 
                      warning=FALSE, 
                      message = FALSE)
```

```{r, include=FALSE}
# PACKAGES-----------------------------------------------
# Tutorial packages
library(vembedr)
library(skimr)
library(yarrr)
library(RColorBrewer)
library(GGally) 
library(tidyverse)
library(plotly)
library(readxl)
library(rvest)
library(biscale)
library(tidycensus)
library(cowplot)
library(units)
```

<br><br>

This tutorial covers reading in every type of file I can think of.

<br>

# 1. General pointers

Here's the structure of EVERY command in R.

```{r, eval=FALSE}

name_youre_saving_it_as  <- COMMAND( "InputFilename.Extension" , other_options)

```

-   **`COMMAND`** is the function used to read the data
    -   e.g.,**`read.csv`** for CSV files <br><br>
-   **`"InputFilename.Extension"`** should be replaced with your file
    name. <br><br>
-   **`other_options`** includes (optional) additional options for your
    command
    -   e.g. nrow specifies the number of rows within the read.csv()
        command <br><br>
-   **`"name_youre_saving_it_as"`** is the name of the variable you
    choose to assign your data to.
    -   e.g `x <- 5` assigns the number 5 to a variable called x
        <br><br>

Using this format, here's an example of how to read the first two rows
of a .csv file called "myfrost.csv," save it as the variable
"froooostdata," and summarize it:

```{r, eval=FALSE}
froooostdata  <- read.csv( "myfrost.csv" , nrow=2)
summary(froooostdata)
```

<details>

<summary>[**Didn't work? Click here to see a few common
mistakes/issues**]{.underline}</summary>

<br>

[**Mistake 1**]{.underline}

```{r, eval=FALSE}
#Incorrect
summary("myfrost.csv")
```

This will summarize the WORD "myfrost.csv" as a character string. To
summarize data, you must first load it into R and assign it to a
variable using \<-.

------------------------------------------------------------------------

[**Mistake 2**]{.underline}

```{r, eval=FALSE}
#Incorrect
read.csv("myfrost.csv")
summary(myfrost)
```

The first line loads the data into R but doesn't save it as a variable,
so it will print the data on the screen. The second line then attempts
to summarize a non-existent variable named myfrost, which will crash.

------------------------------------------------------------------------

</details>

<br><br>

# 2. Organizing your data

You should be running a PROJECT. If so:

-   either place your files in your main project folder.
-   OR create a sub-folder and put your data in there.

If your files are in the main folder, you do not need to use setwd() or
add an address. Commands like this will work.

```{r, eval=FALSE}
mydata <- read.csv("filename.csv")
mydata <- st_read("filename.shp")
mydata <- read.table("filename.txt")
```

If you use a sub-folder, you need to add the location of the sub-folder
to your command. Lets imagine my folder is called Dbata (mis-spelled so
you can see there's nothing special about the name), I would adjust the
commands to look like this

```{r, eval=FALSE}
mydata <- read.csv("./Dbata/filename.csv")
mydata <- st_read("./Dbata/filename.shp")
mydata <- read.table("./Dbata/filename.txt")
```

The "." means "look inside the current project folder", then the "/Data"
means look for a subfolder called Dbata.

------------------------------------------------------------------------

<br><br>

# 3. Read in common data types

## 3.1.  .csv files

```{r, eval=FALSE}
 mydata  <- read.csv("filename.csv")
```

or

```{r, eval=FALSE}
 library(tidyverse) 
 mydata  <- read_csv("filename.csv")
```

<details>

<summary>[**Click here to see a tutorial on these
commands**]{.underline}</summary>

<br>

There are two ways of doing this:

<br>

`read.csv()` : If your file is small, then this command works without
loading any packages at all.

```{r, eval=FALSE}
mydata <- read.csv("filename.csv")
```

For example, if I had a dataset called trees in a subfolder called data,
I would run

```{r, eval=FALSE}
trees.data <- read.csv("./data/trees.csv")
```

------------------------------------------------------------------------

<br><br>

`read_csv()` : This is a more sophisticated way to read in csv data - it
is much faster for large datasets. You first need to add
`library(tidyverse)` to your library code chunk at the top of your
script and run:

```{r, eval=FALSE}
library(tidyverse) # put this in your library code chunk at the top
mydata <- read.csv("filename.csv")
```

For example,

```{r, eval=FALSE}
trees.data <- read_csv("trees.csv")
```

<br><br>

</details>

<br><br>

## 3.2.  .txt files

```{r, eval=FALSE}
 mydata  <- read.txt("filename.txt")
```

or

```{r, eval=FALSE}
 library(tidyverse) 
 mydata  <- read_lines("filename.txt")
```

<details>

<summary>[**Click here to see a tutorial on these
commands**]{.underline}</summary>

<br>

<br><br>

</details>

<br><br>

## 3.3.  .xlsx and .xls files

```{r, eval=FALSE}
 library(readxl) 
 mydata  <- read_excel("filename.xlsx")
```

<details>

<summary>[**Click here to see a tutorial on this
command**]{.underline}</summary>

<br>

<br><br>

</details>

<br><br>

## 3.3. .shp Shape Files

```{r, eval=FALSE}
 library(sf) 
 mydata  <- st_read("filename.shp")
```

IMPORTANT! For this to work, there should be MANY sub-files with the
same name but different extension in your folder (e.g. filename.shp,
filename.dbx..). You need to put ALL of them in your folder, then this
single command reads them in.

<br>

<details>

<summary>[**Click here to see a tutorial on this
command**]{.underline}</summary>

<br>

<br><br>

</details>

<br><br>

## 3.4.  .geo

```{r, eval=FALSE}
 library(sf) 
 mydata  <- st_read("filename.shp")
```

IMPORTANT! For this to work, there should be MANY sub-files with the
same name but different extension in your folder (e.g. filename.shp,
filename.dbx..). You need to put ALL of them in your folder, then this
single command reads them in.

<br>

<details>

<summary>[**Click here to see a tutorial on this
command**]{.underline}</summary>

<br>

<br><br>

</details>

<br><br>

<br><br><br><br><br><br><br><br><br><br><br><br><br><br><br>

<br><br><br>

```{r, eval=FALSE}

# FILE| .csv 
#______________________________________
 mydata  <- read.csv("filename.csv")
 # or
 library(tidyverse) 
 mydata  <- read_csv("filename.csv")

 
# FILE| .txt (tab separated)
#______________________________________
 mydata  <- read.txt("filename.txt",sep="\t", header=TRUE)


# FILE| .xls or xlsx
#______________________________________
 library(readxl) 
 mydata  <- read_excel("filename.xlsx")

 
# FILE| Shape-files.  
# There should be MANY sub-files with the same name 
# but different extension in your folder (e.g. filename.shp, filename.dbx..)
# You need ALL of them in your folder, then this single command reads them in
#______________________________________
  library(sf) 
  mydata  <- st_read("filename.shp")
  
  
  
  
```
