---
title: "Worked Example - A Tourism Focused Analysis of Iowan Farmers Markets"
author: "Dr Helen Greatrex"
output: 
  html_document:
    code_folding: show
    theme:
      color-contrast-warnings: false
      bg: "white"
      fg: "#97a4b8"
      primary: "black"
      secondary: "#2a5736"
      base_font:
        google: Prompt
      heading_font:
        google: Proza Libre
---



# IMPORTANT

This report is the level I would expect from a graduate student with
this assignment, but I want you to aim high. The entire thing from start to end took me 3 hrs, so it's also possible for you by the end of the
semester.\

**WANT TO DOWNLOAD THE UNDERLYING CODE?  CLICK HERE:**


```{r setup, include=FALSE,echo=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message = FALSE,dpi = 300, dev = "svg")

# invisible data read
library(tidyverse)
library(sp)
library(sf)
library(readxl)
library(skimr)
library(tmap)
library(viridis)
library(usmap)
library(rnaturalearth)
library(ggpubr)
library(gridExtra)
library(ceramic)
library(leaflet)
library(leaflet.extras)
library(leaflet.providers)
library(raster)
library(ggmap)
library(dismo)
library(maps)


as.Date_origin <- function(x){
  format.Date(as.Date(x, origin = '1970-01-01'), format = '%b-%d')
}

Sys.setenv(MAPBOX_API_KEY="pk.eyJ1IjoiaGdyZWF0cmV4IiwiYSI6ImNrbmdpNDN6djBkZTQydXBhNDc1enJjZnkifQ.xQWYvOFK5BKQoNWoOacDgA")


```

# Background R Set-Up

*Note, I would normally hide this entire section, using code chunk
option include=FALSE so that so they are invisible. I want you to keep
them in, so I can see your code.*

```{r, results=FALSE,warning=FALSE,message=FALSE}
# Stop warnings printing
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message = FALSE)
```

```{r}
# Libraries
library(sp)
library(sf)
library(tidyverse)
library(tmap)
library(readxl)
library(skimr)
library(spatstat)
```

To see how I came up with this specific code, see the background
workings.

```{r}
# INITIAL CODE
# clear the workspace
rm(list=ls())

# Read in the data
IA_market   <- readxl::read_excel("./Data/Farmers_Markets_Iowa.xlsx",na="NA")

# Remove unreal longitudes
IA_market <- IA_market[which(IA_market$X > -990),]

# Remvbe NAs
IA_market <- IA_market[complete.cases(IA_market$X), ]
IA_market <- IA_market[complete.cases(IA_market$Y), ]

# Remove that one point outside Iowa
IA_market <- IA_market[which(IA_market$X < -90),]

# Choose Columns
IA_market <- IA_market[,c("X","Y","FID","City","County", "Market_Name","Weekday")]

# Make family/grouped/categorical variables into factors
IA_market$City <- as.factor(IA_market$City)
IA_market$County <- as.factor(IA_market$County)
IA_market$Weekday <- as.factor(IA_market$Weekday)

# Rename
names(IA_market)[1] <- "Longitude" 
names(IA_market)[2] <- "Latitude" 

# Make spatial
IA_market.sf <- st_as_sf(IA_market,coords=c("Longitude","Latitude"), crs=4326)

# Make a UTM version
IA_market.sf.utm <- st_transform(IA_market.sf,3744)

# Make a ppp version
IA_market.ppp <- as.ppp(IA_market.sf.utm)

```

# Introduction

[Intended Audience:]{.underline} Matt Kodis, Research Analyst at Iowa
Tourism (<https://www.traveliowa.com/contact-us/>). I have been asked to
provide a spatial analysis of farmers markets in the State, to support
tourism efforts.

## Background

[![The Des Moines Farmers Market, Iowa. Picture sourced from The
DSMPartnership
https://www.dsmpartnership.com/downtowndsmusa/news-blog/blog/how-to-navigate-the-downtown-farmers-market-in-2022](Figures/Report_Fig1_Market.jpeg)](https://www.dsmpartnership.com/downtowndsmusa/news-blog/blog/how-to-navigate-the-downtown-farmers-market-in-2022)

There are over 200 farmers markets in Iowa, providing direct marketing
outlets for fresh locally grown produce; baked goods; and a variety of
products from eggs to meat, wine, cheese and crafts.

Understanding the spatial pattern of farmers markets is crucial for
enhancing our tourism strategies in Iowa. It allows us to make informed
decisions that can significantly impact our marketing efforts and the
overall visitor experience. First and foremost, having this knowledge
enables us to effectively target and engage potential tourists. We can
tailor our promotional materials and itineraries to highlight the
farmers markets that align with their interests and preferences. It's
not just about quantity; it's about showcasing the right markets in the
right places to attract the right audience.

There has been some research on this topic in other areas of the USA.
For example, Tong et al[^1] (2012) theoretically modeled the placement
of new markets in Arizona; Mack and Tong[^2] (2015) used surveys to
characterize the spatial and temporal patterns of farmers' market
visits; and Sage[^3] (2012) used point pattern analysis to consider the
pattern of market locations in Washington State.

[^1]: Tong, D., Ren, F., & Mack, J. (2012). Locating farmers' markets
    with an incorporation of spatio-temporal variation. Socio-Economic
    Planning Sciences, 46(2), 149-156.
    <https://www.sciencedirect.com/science/article/pii/S0038012111000346>

[^2]: Mack, J., & Tong, D. (2015). Characterizing the spatial and
    temporal patterns of farmers' market visits. Applied Geography, 63,
    43-54.
    <https://www.sciencedirect.com/science/article/pii/S0143622815001496>

[^3]: Sage, J. L. (2012). A geographic exploration of the social and
    economic sustainability of farmers' markets and the rural
    communities that make them work. Washington State University. PhD
    Thesis <https://www.proquest.com/docview/1032783987>

## Dataset

The specific dataset I am using for this analysis is obtained from the
Iowa State GIS portal[^4], of 175 markets in 2018. This reported that
the original source of the data was the Iowa Department of Agriculture
and Land Stewardship Data Directory
([https://www.idalsdata.org/](https://www.idalsdata.org/fmnp/index.cfm?fuseaction=main.formFarmersMarketDirectory).)),
which contains the details of all licensed farmers markets currently in
Iowa. The data is collected from license records, so should be complete
across the state.

[^4]: <https://geodata.iowa.gov/maps/farmers-markets>

Importantly, the GIS portal was last updated on 15 November 2018 - so
the data I am analyzing is not fully up to date. For example, my
analysis will not reflect recent updates or closures due to the COVID
pandemic. Given time for a more detailed analysis, I would pull the most
recent data directly from IDALS.

The data will also not cover markets in other States that are just on
the border with Iowa. This means we might get 'edge effects' in our
spatial analysis. For example, as shown in the Figure below - there are
towns to the West of Iowa that spread across the border. It would be
reasonable to assume there were more than 2 markets in this town, but we
only have data for the Iowan side.

[![Edge effects in Iowan Markets. Map annotated, with the original
obtained from from
https://geodata.iowa.gov/datasets/iowa::farmers-markets/](Figures/Report_Fig2_MarketEdge.png)](https://geodata.iowa.gov/datasets/iowa::farmers-markets/)

### Structure of the data

As shown below in the code, there are 165 markets available for study
after missing data was taken into account. The location is stored in
Lat/Long format and the data has these variables available for study.

-   `Longitude` : The longitude of each market in degrees

-   `Latitude` : The latitude of each market in degrees

-   `FID` : This is the unique ID of each market applied by the Iowa GIS
    portal

-   `City` : The name of the city the market is held in

-   `County` : The name of the Iowa county the market is held in

-   `Market_Name` : The name of the market

-   `Weekday` : The day of the week that the market is held on

<br><br>

## Quality control/Data wrangling

Some quality control was carried out before the data was analysed:

1.  A few markets had corrupted or missing locations. These were
    removed. <br>
2.  Only relevant variables were kept for analysis. The columns
    *Open_Hours*, *Open_Dates* and *Location* were removed but could be
    analysed at a later date. <br>
3.   The *County*, *Weekday*, and *City* columns were converted to
    factor to aid analysis.<br>
4.  The data was converted to a spatial version using the package sf and
    the UTM map projection for Iowa was applied. A spatial version of
    the data for point package analysis (IA_market.ppp) was created.<br>

<br><br>

# Exploratory analysis
