---
title: "Lab 3: Bei Data Set Spatial Analysis"
date: "`r Sys.Date()`"
author: "dpd5576"
output:
  rmdformats::downcute:
    self_contained: true
    default_style: "dark"
    downcute_theme: "default"
---
was thinking about doing a spatial analysis project themselves.

# Lab 3B Farmers Markets

### Reading in the data

```{r}
market    <- read_csv("./Data/Farmers_Markets_NYShort.csv")
```

### Making the data spatial and utm

```{r}
#the first like makes an "sf" version of the data that helps us make maps
market.sf <-st_as_sf(market,coords=c("Longitude","Latitude"),crs=4326)

#changing the map proection from long/lat to "UTM"
market.utm <-st_transform(market.sf,crs=26918)
```

```{r}
#making a map!
tmap_mode("view")
qtm(market.utm)
```

### Making the data ppp/spatstat format

```{r}
#this reads the data into a ppp format, and then sets our study area to the box surrounding the points
market.ppp <- as.ppp(market.utm)
plot(market.ppp,use.marks = F, cex = 1, pch = 4)
```

## Challenge A

1.  **Summarize this data set. What does it contain?**

    (Answer)

2.  **What spatial processes do you see so far. Are there 1st and/or 2nd order autocorrelation processes that might be behind them?**

    (Answer)

## Challenge B

1.  **Nearest neighbor plots:**

    ```{r}
    #nearest neighbor analysis
    market.ppp <- nndist(market)
    #this command then makes the histogram
    hist(market.ppp,br=50)
    ```

    (explain code and results)

2.  **Average nearest neighbor**

    ```{r}
    paste( c("The average nearest neighbor of my Market data is:"),
          round(mean(market.ppp),3),"m away")
    ```

    (explain code and results)

3.  **Calculating the Clarke-Evans Ratio**

    ```{r}
    ?clarkevans
    ```

    ```{r}
    market.obs.RRatio <- clarkevans(market.ppp)
    market.obs.RRatio
    ```

    (explain code and results)

4.  **Conducting hypothesis test on Clarke-Evans Ratio Output:**

    ```{r}
    clarkevans.test(market.ppp)
    ```

    (NOTE: H0 and H1 taken directly from Lab instructions)

    -   H0: The pattern shows complete spatial randomness caused by an independent random process. R=1. E.g. If we simulated 100 random point patterns and calculated the Clarke Evans ratio for each, we would get a histogram with a mean of 1.

    -   H1: The pattern is unusually DIFFERENT (e.g. strangely clustered OR strangely uniform) compared to what we would expect in H0. E.g. R would be high or low.

    ```{r}

    ```

5.  **General Interpretation of the data**

    (answer)

6.  **One-Sided Clarke Evans Hypothesis Test**

    (answer)

7.  Fallacy Questions

    -   [What are edge effects in point pattern data? Are they going to be an issue for this specific dataset/analysis and why?]{.underline}

        (Answer)

    -   [What is the locational fallacy? Is it going going be an issue for this SPECIFIC dataset/analysis and why?]{.underline}

        (Answer)

    -   [What is MAUP? Is it going be an issue for this SPECIFIC dataset/analysis and why?]{.underline}

        (Answer)

## Challenge C

1.  **Ripley's L-Function**

    ```{r}
    market.lest <- Lest(market,correction = "Ripley")
    #creating a collection of simulated point patterns based on our data, but the points have been rearrange using independent random process (for each one, we calculate the L function, lest)

    market.lest.IRsimulation <- envelop(market,
                                        Lest, correction = "Ripley",
                                        verbose = F,
                                        nsim=100, nrank=1,
                                        savepatterns = FALSE, savefuns = FALSE)
    ```
